Problems with the characterisation of accreting Brown Dwarfs: A public grid of simulated SEDs and photometry

### 2010-05-19 BD_disc/temp.tex
Shortened Rotation rate section (parameters) and changed inner_temp.ps
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@2272 25464f21-6099-402c-b30a-640f1d40d38b

% mn2esample.tex
%
% v2.1 released 22nd May 2002 (G. Hutton)
%
% The mnsample.tex file has been amended to highlight
% the proper use of LaTeX2e code with the class file
% and using natbib cross-referencing. These changes
% do not reflect the original paper by A. V. Raveendran.
%
% Previous versions of this sample document were
% compatible with the LaTeX 2.09 style file mn.sty'
% v1.2 released 5th September 1994 (M. Reed)
% v1.1 released 18th July 1994
% v1.0 released 28th January 1994

\documentclass[useAMS,usenatbib]{mn2e}

% If your system does not have the AMS fonts version 2.0 installed, then
% remove the useAMS option.  
% useAMS allows you to obtain upright Greek characters.
% e.g. \umu, \upi etc.  See the section on "Upright Greek characters" in
% this guide for further information.
%
% If you are using AMS 2.0 fonts, bold math letters/symbols are available
% at a larger range of sizes for NFSS release 1 and 2 (using \boldmath or
% preferably \bmath).
%
% The usenatbib command allows the use of Patrick Daly's natbib.sty for
% cross-referencing.
%
% If you wish to typeset the paper in Times font (if you do not have the
% PostScript Type 1 Computer Modern fonts you will need to do this to get
% smoother fonts in a PDF file) then uncomment the next line
% \usepackage{Times}
\usepackage{lscape}
\usepackage{graphicx}

%%%%% AUTHORS - PLACE YOUR OWN MACROS HERE %%%%%
\newcommand\mnras{MNRAS}
\newcommand\pasp {PASP}
\newcommand\aj   {AJ}
\newcommand\apj  {ApJ}
\newcommand\apjl  {ApJL}
\newcommand\apjs  {ApJS}
\newcommand\aap  {A\&A}
\newcommand\aaps  {A\&AS}
\newcommand\apss  {AP\&SS}
\newcommand\gca  {Geochim. Cosmochim. Acta}
\newcommand\araa {ARA\&A}
\usepackage{multirow}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title[Accreting Brown Dwarfs]{Problems with the characterisation of
  accreting Brown Dwarfs: A public grid of simulated SEDs and
  photometry}  \author[N.J.  Mayne and Tim
Harries]{N.J. Mayne$^{1}$\thanks{E-mail: nathan@astro.ex.ac.uk
    (NJM)} and Tim Harries$^{1}$\\
  $^{1}$ School of Physics,
  University of Exeter, Stocker Road, Exeter, EX4 4QL.\\
}
 \begin{document}

\date{Accepted ?. Received ?; in
  original form ?}

\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2009}

\maketitle

\label{firstpage}

\begin{abstract}

   1) All graphs of SEDS in log log units.

  1a) Add in the new data and write in using more masses.

  1b) Write in the extra inclinations-check from incs.lis.

  1c) Write in that using same points for photosphere.dat

  1) BOUY ET AL 2008, NEUGEBAUER 1984 AND SKRUTSKIE 2006 TRUNCATED AUTHOR LIST.

  2) CHANGE SYMBOLS OF NAKED STARS IN IR EXCESS AND IR EXCESS 2 TO
  MAKE CLEARER.

  3) REWRITE CONCLUSIONS WITH MORE REFERENCE TO
  \cite{2004MNRAS.351..607W}, \cite{2006A&A...452..245N} AND
  \cite{2006MNRAS.370L..10C}.

  4) REDO \& CHECK ALL FIGURES.

  5) COMPARE FINAL INNER RADIUS, AFTER SUBLIMATION, TO COROTATION
  RADIUS-HOT GAS STUFF.

  6) ADD IN DISCUSSION FOR MIPS AND CHECK ZEROPOINTS

  7) NEED A DISCUSSION OF THE DISC SCALEHEIGHT ETC (PROBABLY IN MODEL
  OR PAR SPACE): $\rho \propto R^{-\alpha}$, $\alpha =
  2.25$. $h\propto R^{\beta}$, $\beta = 1.25$. $\sigma = R^{(\beta
    -\alpha)}$, AND IS CONSERVED. $R_{\rm outer}=300$ AU. DISC SCALEHEIGHT
  AT 100 AU $H(100)=12.5$ AU

  8) ADD IN DISCUSSION OF THE TIME TAKEN FOR EACH MODEL AND
  SED-JUSTIFY SMALL NUMBER OF GRID. EACH MODEL (FOR ALL INCLINATIONS)
  =** HOURS OF SUPERCOMPUTER TIME ON ** CORES

  9) WHAT ABOUT EFFECT OF VERTICAL WALL FOR LONGER PERIOD OBJECTS (NO
  DUST SUBLIMATION) VERSUS SHORTER PERIOD AND CURVED WALL

  10) PUT IN INNER BOUNDARY PICTURES

  11) SORT OUT PLACEMENT OF CLEARPAGE.

  12) SORT RESULTS BIT ABOUT IRAC FLUX, THIS IS FROM REGIONS BEHIND
  THE INNER EDGE SO WHY IS THE TREND REVERSED, IT IS NOT COMING FROM
  THE OUTER FLARED REGIONS OF THE DISC BUT FROM JUST BEHIND THE PUFFED
  UP INNER EGDE. THEREFORE, FOR LONGER PERIOD OBJECTS THE INNER EDGE
  IS SMALLER AND THE DISC FLARED SO REGIONS JUST BEHIND CAN EMIT MORE
  FLUX. FOR SHORTER PERIOD OBJECTS, THE INNER EDGE IS PUFFED UP AND
  DOMINATES THE SED.

  13) TO PROVE THE POINT ABOVE PRODUCE AN IMAGE BATCH  FILE TO PRODUCE
  IMAGES FOR A TEST CASE I.E. CLOSE INNER EDGE DISTANT ONE ETC.

  14) Add in reference to the Mike Meyer paper about changes due to
  inner radius.

15) Reference to bouy et al 2008 showing that the outer radius has a
negligible effect

16) IMAGES.

17) Add in the Herschel filters.

18) Disc mass fraction.

19) IMF implications.

20) ADD IN RHO OPH 102 STUFF- and add in comparison to stars without accretion

21) Write in a bit about do we expect to see the correlation with
rotation rate.

22) Colours of 0.01Msol and 0.01Gyr models are very different.-REMOVE

23) Notes abnout isong 200 pts for SEDS naked in consisitency checks
used full resolution

24) NOTE ABOUT THE FACT THAT 0.01Msol STARS ARE OFF THE BOTTOM OF SOME
PLOTS AS THEY DIVE DOWN VERY ABRUPTLY.ASK ISABELLE ABOUT THIS

25) Add in comments to mention that due to numerics some models have
inner edges up to 100K above the 1700K mean dust sublimation temps.

26) Note about colour colour plots cut in magnitude for J and removed
the 0.01Msol stars from the naked

27) For detections add in piece about the overlapping stars i.e what
would be missed.

28) missing 0.01 in alpha graph also.

29) Adjust size of the smaller figures (i.e non-four) to fit.

30) Change argument around to say rotation period correlation not
present.

31) Plots of inner temp against dust sublimation inner radius etc

32) inclination against magnitude for different accretion rates.-one mass

33) Sot out JJK isochrone

34) Note about r=300 but outer radius doesnt matter largey see bouy
and vitell 2008 or something.

   \end{abstract}

\begin{keywords}
  stars:evolution -- stars:formation -- stars: pre-main-sequence --
  techniques: photometric -- catalogues -- (stars) Hertzsprung-Russell
  H-R diagram
\end{keywords}

\section{Results: Analysis and Discussion}
\label{results}

\textbf{Intro-Questions to answer:IMF, Formation, Disc timescales and
  $\dot{M}\propto M_*^{\approx 2}$=Ages, Masses and Disc fractions}

Uncertainty remains over the formation mechanisms, resulting IMFs and
subsequent interaction mechanisms with associated circumstellar disc,
for BDs. Furthermore, no current theoretical explanation can be
demonstrated for the suspected stellar mass-accretion rate
relationship \citep[$\dot{M}\propto
M_*^2$,][]{2003ApJ...592..266M,2004A&A...424..603N,2006A&A...452..245N}
derived from observations and doubt remains over the existence of a
disc-locking mechanism in BDD systems. Theoretical models and are
built upon constraints, such as ages, masses and disc fractions, and
correlations, such as the accretion rate to stellar mass relation,
derived from the data. Critically, these resulting constraints and
relationships involve both the application of existing models, such as
isochrones to derive ages, and the selection of bias-minimised
samples, for instance removal of foreground or background contaminant
stars. Clearly, if one is to adopt a new theoretical model or use an
observed correlation as a constraint one must explore the suspected
intrinsic biases and the reliability of the selection criteria.

In this Section we explore the effects of disc presence and accretion
(across the described representative ranges, Section \ref{par_space})
for our model grid. Essentially the premise is that if one accepts the
assumptions and physical processes within our model grid would one be
able to derive robust parameters using existing methods and
furthermore, are the correlations found in the data intrinsic to the
grid.

Practically, most parameters for young pre-MS stars are derived from
surveys of populations, usually open clusters, using broadband
photometry and subsequently constructed colour-magnitude and
colour-colour diagrams (CMDs and CoCoDs respectively,
hereafter). Therefore, in this Section, firstly we demonstrate the
prohibitive effects on the broadband photometry of varying our input
parameters (Section \ref{photometry}). Secondly, we explore the
underlying causes of the photometric scatters and intrinsic
correlations using the SEDs and structure of individual BDD systems
(Section \ref{sed_causes}).

\subsection{Photometry}
\label{photometry}

\subsubsection{Ages and Masses}
\label{age_mass}

For the derivation of ages optical CMDs, in particular in \textit{V,
  V-I}, are most often used, and indeed most suitable. Whereas, IR
CMDs, such as a \textit{J, J-K} CMD, are most suitable for mass
derivation \cite[see references and discussions in][ and Section
\ref{derived}]{2007MNRAS.375.1220M,2008MNRAS.386..261M}.

To delineate the effects of the accretion rate and circumstellar discs
we have subdivided the grid into two groups, those with accretion
rates typical for higher mass CTTS objects, defined as
$\dot{M}=$10$^{-12}M_{\odot}yr^{-1}$ (negligible) to
$\dot{M}=$10$^{-9}M_{\odot}yr^{-1}$ and those with elevated accretion
rates, where $\dot{M}>$10$^{-9}M_{\odot}yr^{-1}$. For several of the
plots in this section the magnitude and colours for the
$M_*=0.01M_{\odot}$ at high inclinations become extremely faint and
red, respectively. In some cases these objects appear slightly without
the scale of the diagram the axes where limited in this way to be able
to show the changes in colour an magnitude for the majority of stars
better. The colours and magnitudes for these lowest mass stars are not
necessarily unreliable but simple hinder the aesthetics of the plots,
and as they are at the limits of our grid we have decided to omit them
from some Figures by trimming the axes. Additionally, stars at the
highest inclinations, i.e edge on disc systems, are often omitted from
the Figures due to their extremely faint magnitudes, meaning they
would not be practically observable.

\textit{Derivation of Ages}

Figure \ref{spread_opt_VVI} shows four CMDs in \textit{V, V-I} with
isochrones constructed from our data (comprising all masses from Table
\ref{par_space_table}, excluding $M_*=0.01M_{\odot}$ for 10 Myr) at
ages of 1 Myr and 10 Myrs, solid and dashed lines respectively. These
isochrones are produced using the naked BD photometry at an accretion
rate of -12 log$(\frac{\dot{M}}{M_{\odot}}yr^{-1})$ and a areal spot
coverage of 10\%. The solid green line in the \textit{bottom left
  panel} only is the 1Myr isochrone for naked systems at an accretion
rate of -9 log$(\frac{\dot{M}}{M_{\odot}}yr^{-1})$ and a areal spot
coverage of 10\%. The solid green line in the \textit{top left panel}
only is the 1 Myr from \cite{2000A&A...358..593S} adjusted to a
distance of 250pc and an extinction of $A_V=2$ mag, simulating a
background population of CTTS stars. The remaining dots are the
simulated photometry for the BD disc systems with accretion rates
classed as typical (-9 to -12 log$(\frac{\dot{M}}{M_{\odot}}yr^{-1})$)
at 1 Myr (only) for all areal coverages, periods and inclinations (see
Table \ref{par_space_table}). The four panels then highlight the
scatter in CMD space caused by variation across our input parameter
space. The \textit{top left panel} splits the accretion rates into
three, -11 \& -12 log$(\frac{\dot{M}}{M_{\odot}}yr^{-1})$) as blue
dots, -10 log$(\frac{\dot{M}}{M_{\odot}}yr^{-1})$ as black dots and -9
log$(\frac{\dot{M}}{M_{\odot}}yr^{-1})$ as red dots. The
\textit{bottom left panel} shows the scatter due to period changes,
with systems with periods of 5 days shown as blue dots and 0.5 days as
red dots. The \textit{top right panel} shows the scatter caused by
changes in areal coverage, with blue dots showing photometry from 1\%
coverage systems and red dots for 10\% coverage systems. Finally, the
\textit{bottom right panel} shows the results separated by inclination
angles. The inclinations have been separated into three groups for the
purposes of CMD analysis. These groups are $\theta \leq 48^{\circ}$ as
blue dots (classed as face-on systems), $\theta> 56$ \& $64^{\circ}$
(classed as the expected systems, as the expectation value of
$cos(\theta)=60^{\circ}$) as black dots and $\theta \geq 71^{\circ}$
(classed as edge-on systems) as red dots.

\begin{figure*}
  \vspace*{348pt}
  \special{psfile="Fig/spread_opt_VVI.ps"
   hoffset=470 voffset=0 hscale=60 vscale=60 angle=90}
 \caption{CMDs in $M_V, (V-I)_0$ showing,as lines, the isochrones for
   all masses included in Table \ref{par_space_table} for ages 1 Myr
   (solid) and 10 Myr (dashed). The isochrones are for naked BDs
   without a disc, for the accretion rate -12
   log$(\frac{\dot{M}}{M_{\odot}}yr^{-1})$ (shown as in black), with
   an areal coverage of 10\%. The solid green line in the \textit{top
     left panel} only is the 1 Myr from \citet{2000A&A...358..593S}
   adjusted to a distance of 250pc and an extinction of $A_V=2$ mag,
   simulating a background population of CTTS stars. The dots are the
   simulated photometry, of the star and disc system, for all models
   at an age of 1 Myr with the accretion rates classed as typical (-9
   to -12 log$(\frac{\dot{M}}{M_{\odot}}yr^{-1})$). \textit{Top left
     panel} shows the accretion rates separated into three groups -12
   \& -11, -10 and -9 log$(\frac{\dot{M}}{M_{\odot}}yr^{-1})$ as blue,
   black and red dots respectively. \textit{Bottom left panel} shows
   BDD systems with a rotation period of 5 days as blue dots and 0.5
   days as red dots. \textit{Top right panel} shows those simulations
   with an areal coverage of 1\% as blue dots and 10\% as red
   dots. Finally, the \textit{bottom right panel} shows the systems in
   three groups of inclinations, $\theta \leq 48^{\circ}$, as blue
   dots, $\theta \geq 71^{\circ}$ as red dots and $\theta = 56$ \&
   $64^{\circ}$ as black dots. \label{spread_opt_VVI}}
\end{figure*}

As can be seen in Figure \ref{spread_opt_VVI} current accretion in a
star and disc system, for rates typical in the BD mass regime, creates
a scatter in our simulated photometry indicative of a much larger
isochronal age spread than 10 Myr. Indeed, for many BDD systems even
at nominal accretion rates of -11 or -12
log$(\frac{\dot{M}}{M_{\odot}}yr^{-1})$, for our simulations, the
colours of these stars move significantly blueward of the expected BD
locus in a\textit{V, V-I} CMD. As the dots in Figure
\ref{spread_opt_VVI} are the simulated photometry of BDD systems over
a range of typical input parameters (see Section \ref{par_space}), an
observed coeval 1 Myr population could reasonably be expected to show
a similar scatter. The naked isochrones show that this spread in
simulated photometry, would lead to a spread in isochronal age greater
than the $\approx$10 Myr spread claimed for higher mass stars in some
star forming regions (for instance the ONC). Furthermore, for the
higher accretion rates of -9 or -10
log$(\frac{\dot{M}}{M_{\odot}}yr^{-1})$ the movement of the star
within the CMD will effectively move the star into the contamination
region expected for background CTTS or MS stars at a \textit{V-I} of
$\leq$1.5 and $M_V 12-10$, and as such the star would not be included
in a photometrically selected BD sample. The solid green line, in the
\textit{top left panel}, showing the 1 Myr isochrone of
\cite{2000A&A...358..593S} at a distance of 250 pc and extinction of
$A_V=2$ mags shows that the BDD systems with higher accretion rates
could easily be confused for a background CTTS or MS population (as MS
stars are simply less luminous at a roughly constant \textit{V-I} than
the CTTS counterparts). Furthermore, scatter, although somewhat
reduced, can be observed in the simulated photometry in the negligibly
accreting systems, for example in the \textit{top left panel} some red
dots, accreting at -12 or -11 log$(\frac{\dot{M}}{M_{\odot}}yr^{-1})$
(which will be negligible compared to the photospheric flux, see
Figure \ref{acc_dom}), still scatter below the 10 Myr. These objects
may still be included in a `wide' photometric selection. This scatter
is a strong function of inclination, as shown in the \textit{bottom
  right panel}, where, as the inclination is increased the objects are
pushed lower in the CMD. Indeed, for the edge on cases some objects
have magnitudes fainter than show in Figure \ref{spread_opt_VVI}
($M_V\approx 30$). This is expected as the star becomes obscured by
the flared disc, interestingly for these systems the \textit{bottom
  right panel} shows that this occurs for inclinations above around
71$^{\circ}$ in most cases yet even for the lower inclination angles
some objects have very faint magnitudes, this is due to the disc
flaring leading to a smaller opening angle and is discussed later in
Section \ref{disc_struct}. The \textit{top right panel} shows that the
scatter, in \textit{V}, in our simulated photometry from the
isochrones is significant for both areal coverages. However,
significant scatter blueward in colour is slightly more evident for
the lower areal coverage and therefore hotter accretion spot
temperature, as is expected given that the higher spot temperature
contributes more flux blueward of the photosphere. Finally, the
\textit{bottom left panel}, shows significant scatter in the simulated
photometry for both adopted rotation periods. This panel also shows
that scatter in this diagram is not obviously correlated with the
rotation period and therefore inner edge position. Additionally, the
\textit{bottom left panel} shows that for naked systems accretion
rates of -9 log$(\frac{\dot{M}}{M_{\odot}}yr^{-1})$ lead to the stars
being scattered into the region occupied by background CTTS and
contamination (the lower accretion rates are all coincident with the
-12 log$(\frac{\dot{M}}{M_{\odot}}yr^{-1})$ for 1 Myr). Overall, as
found in \cite{2004MNRAS.351..607W}, the dominant scattering effect
for BD disc systems in an optical CMD appears to be caused by
inclination and therefore obscuration effects of the disc on the
star. This suggests that for a given photometric survey of BDD systems
accreting at typical accretion rates and with an expected range of
inclinations (centred on around 60$^{\circ}$) one would expect to
exclude a significant fraction of these objects from any isochrone
based selection.

Similar scattering occurs when using CMDs constructed using \textit{R,
  R-I} \citep[using the photometric system
of][]{1998A&A...333..231B}. Figure \ref{spread_opt_RRI} shows the
\textit{R, R-I} CMD.

\begin{figure*}
  \vspace*{348pt}
  \special{psfile="Fig/spread_opt_RRI.ps"
   hoffset=470 voffset=0 hscale=60 vscale=60 angle=90}
 \caption{As for Figure \ref{spread_opt_VVI} but the 1 Myr isochrone
   of \citet{2000A&A...358..593S} is omitted and the CMDs are for
   $M_R, (R-I)_0$. \label{spread_opt_RRI}}
\end{figure*}

Therefore, if one adopts the range of input parameters we have used
(see Section \ref{par_space} for justification), our simulated
photometry shows, qualitatively, that a coeval 1 Myr population of
accreting BD stars and BDD systems, with typical accretion rates and
range of inclinations, will exhibit a scatter in apparent isochronal
age of $>$10 Myr. Furthermore, some objects will be scattered below
and blueward of a 10 Myr isochrone. This has been shown for CMDs in
\textit{V, V-I} and \textit{R, R-I}. It is important to note at this
point that these conclusions are qualitative, and obviously based on
our assumptions. However, the repercussion for isochronal age
derivation and sample selection in the BD regime could be
profound. Indeed this study has only included the effects of current
or ongoing accretion from a disc, it has totally neglected any effects
of accretion, both past and present, on the evolution of the central
star. This past accretion could also act to reduce the stars radius,
accelerating contraction
\citep{1999MNRAS.310..360T,1999A&A...342..480S} and introducing
additional scatter in a coeval population proportional to the range in
accretion rates \citep[See][for full
discussion]{2008MNRAS.386..261M}. Our findings support those of
\cite{2004MNRAS.351..607W}, showing that a significant scatter is
caused by obscuration of the central source by the central disc. In
addition we have shown that the scatter blueward, into the CTTS or
background MS locus, is caused by active accretion at levels found in
CTTS stars, but currently not regularly recorded in stars classified
or selected as BDs.

\textbf{Derivation of Masses}

Similar consequences are apparent when attempting to derive masses
from isochrones using CMDs. We have constructed CMDs using the
simulated photometry of our model grid in the \textit{JK}
magnitudes. Figure \ref{spread_ir_JJK} shows CMDs in \textit{J, J-K} for
the same data as Figure \ref{spread_opt_VVI}. The symbols have the
same meanings as in Figure \ref{spread_opt_VVI}.

\begin{figure*}
  \vspace*{348pt}
  \special{psfile="Fig/spread_ir_JJK.ps"
   hoffset=470 voffset=0 hscale=60 vscale=60 angle=90}
 \caption{As for Figure \ref{spread_opt_VVI} but the CMDs is for $M_J,
   (J-K)_0$.\label{spread_ir_JJK}}
\end{figure*}

As is shown by Figure \ref{spread_ir_JJK} a similar scatter is
observed in our simulated photometry in the IR CMD as is evident in
its optical counterparts. Once again, as shown in the \textit{top left
  panel}, scatter is seen in both the nominal and higher accretion
rates (-9 to -12 log$(\frac{\dot{M}}{M_{\odot}}yr^{-1})$), meaning
derivation of masses for objects using IR isochronal fitting of these
objects would be very difficult. However, in this CMD the scatter is
generally redward of the isochrone, this is still however into the
region where background CTTS stars would be expected. The \textit{top
  right panel} shows no correlation with areal coverage of the hot
spot. Interestingly the \textit{bottom left panel} shows no clear
evidence for a correlation in CMD position, and therefore distance
from the isochrone, with rotation period of the system as found in
\cite{1997AJ....114..288M}, this is discussed in more detail in
Section \ref{disc_struct}. Additionally, \textit{bottom left panel}
agains shows significant scatter away from the isochrone for the 1 Myr
naked stars at an accretion rate of -9
$(\frac{\dot{M}}{M_{\odot}}yr^{-1})$, however in this case the scatter
is in the opposite sense (blueward as opposed to redward) to that
caused by accretion and disc presence in the BDD systems. Again,
scatter is still caused by changes in inclinations, as shown in the
\textit{bottom right panel}, with fainter magnitudes for more inclined
systems, as expected however, this effect causes a smaller change in
magnitude in the IR due to less of the disc appearing optically thick
at these wavelengths.

Therefore, as for derivations of age, given our assumptions, and the
adopted range of input parameters, a 1 Myr coeval population scatters
significantly in a \textit{J, J-K} CMD. The \textit{J, J-K} CMD is
representative of CMDs constructed from the \textit{JHK}
passbands. So, qualitatively at least, derivation of masses using
isochronal fitting of our simulated photometry would not produce
reliable or accurate results. Again, as discussed for optical CMDs,
these simulations do not include any of the effects of past or current
accretion on the evolution of the central object, which is expected to
significantly exacerbate the observed scatter. This is also true for
the naked systems with small increases in accretion rate leading to
significant scatters in the CMDs. Additionally, given the evidence
discussed in \cite{2004MNRAS.351..607W}, supported by the evidence
from our own simulated photometry shown in Figures
\ref{spread_opt_VVI}, \ref{spread_RI} and \ref{spread_ir_JJK}, it is
clear that young, accreting, BDD systems are likely to be
misidentified when using photometric colours or magnitudes, even at
generally canonical accretion rates. This casts doubt on the veracity
of the mass to accretion rate relationship. For our model grid we have
not assumed any such relation, therefore, as our data would also show
a similar relation it suggests that the observed result may be caused
by intrinsic scattering. The relation, $\dot{M}\propto M_*^{2}$
suggests that their is a dearth of lower mass stars accreting at
higher rates. We have shown that for accretion rates in the range -12
to -9 log$(\frac{\dot{M}}{M_{\odot}}yr^{-1})$ the BDD systems with
higher accretion rates would be preferentially missed using
photometric or isochronal selection.

\subsubsection{Extreme accretion rates}
\label{elevated_accn}

\textbf{CMDs}

BDD and naked BD systems with elevated or extreme accretion rates are
also likely to be miss-classified, or not observable in the same
dynamic magnitude range as a BD survey.  Figure \ref{high_acc_opt}
shows our isochrones for the naked BD systems with accretion rates of
-6, -7 \& -8 log$(\frac{\dot{M}}{M_{\odot}}yr^{-1})$ at 1 Myr (solid
lines in red) with areal coverages of 10\%, and the pre-MS isochrone,
at 1 Myr, of \cite{2000A&A...358..593S} for higher mass stars, down to
$M_*=$0.1$M_{\odot}$ (solid black line). The remaining dots are the
simulated photometry from our model grid for stars with an age of 1
Myr and accretion rates of -6, -7 \& -8
log$(\frac{\dot{M}}{M_{\odot}}yr^{-1})$, and the complete range of all
other input parameters (see Table \ref{par_space_table}. As can be
seen by comparing Figure \ref{spread_opt_VVI} with Figure
\ref{high_acc_opt} the simulated photometry for the naked systems
changes dramatically with increasing accretion rate, meaning that even
without the effects of a disc individual highly accreting BDs
dramatically change position in an optical CMD. The simulated
photometry of the disc systems is then separated in each panel. The
\textit{top left panel} shows systems with accretion rates of -6, -7
and -8 log$(\frac{\dot{M}}{M_{\odot}}yr^{-1})$ as red, black and blue
dots respectively. The \textit{top right panel} distinguishes the
areal coverages of 1 and 10\% as red and blue dots respectively. The
\textit{bottom left panel} separates the two rotational periods of 5
and 0.5 days into blue and red dots respectively. Finally, the
\textit{bottom right panel} plots the three adopted inclinations
groups $\theta \leq 48^{\circ}$, $\theta> 56$ \& $64^{\circ}$,
$cos(\theta)=60^{\circ}$) and $\theta \geq 71^{\circ}$, as blue, black
and red dots respectively.

The \textit{top left panel} shows that these objects are generally
scattered to an apparently higher (isochronal) mass and older
(isochronal) age, with the scatter increasing with accretion
rates. This means that as the accretion rate increases the star is
pushed to bluer and brighter colours and magnitudes, effectively, to
higher stellar masses. The \textit{top right panel} shows that this
effect is apparent for both areal coverages, it is perhaps surprising
that in general there does not appear to be a reduction in the
magnitude of the scatter when one moves to the larger areal
coverage. The \textit{bottom left panel} shows again that there is no
correlation in CMD position with rotation rate, and therefore
co-rotation radius (again this is discussed further in Section
\ref{disc_struct}. Finally, the \textit{bottom right panel} again
shows that the system inclination scatters the resulting position
within the CMD, with greater inclination meaning greater obscuration
of the central star, by the disc, and therefore fainter magnitudes and
redder colours, and it is clear that photometrically a highly
obscured, highly accreting BDD system could be photometrically
classified as contamination or a higher mass CTTS system.

Once again, given our assumptions, qualitatively, young, pre-MS BD
disc systems accreting at higher rates would not be identifiable as
such from optical CMDs and isochronal fitting. Meaning it is likely
that our data would indicate a stellar mass to accretion rate
relation, even though we have not adopted such a relationship.

\begin{figure*}
  \vspace*{348pt}
  \special{psfile="Fig/high_acc_opt.ps"
   hoffset=470 voffset=0 hscale=60 vscale=60 angle=90}
 \caption{The optical CMD, $M_V, (V-I)_0$ showing the effects of
   extreme accretion rates. The solid black line is the pre-MS
   isochrones of \citet{2000A&A...358..593S} at 1 Myr for masses down
   to 0.1 $M_{\odot}$, the solid red lines are our derived isochrones
   for the naked photospheric models at accretion rates of -6, -7 \&
   -8 log$(\frac{\dot{M}}{M_{\odot}}yr^{-1})$, colours of red, black
   and blue respectively, using an areal coverage of 10\% (the coolest
   and therefore lowest accretion luminosity). The dots are the
   simulated photometry for the BDD systems for the three highest
   accretion rates (as for the isochrones). \textit{Top left panel}
   shows the accretion rates separated into three groups -6, -7 and -6
   log$(\frac{\dot{M}}{M_{\odot}}yr^{-1})$ as red, black and blue dots
   respectively. The data in the remaining panels is separated into
   the groups used in Figure \ref{spread_opt_VVI} using the same
   symbols and colours.\label{high_acc_opt}}
\end{figure*}

A similar situation is apparent for the IR passbands, where isochrones
are often used to derive stellar masses. Figure \ref{high_acc_ir} is
the IR version of Figure \ref{high_acc_opt}, the only difference
between the figures are the colours and magnitudes, the conventions,
symbols and models are the same. Figure \ref{high_acc_ir} shows the
CMDs using \textit{J, J-K}, also using the same conventions for each
of the \textit{four panels}.

\begin{figure*}
  \vspace*{348pt}
  \special{psfile="Fig/high_acc_ir.ps"
   hoffset=470 voffset=0 hscale=60 vscale=60 angle=90}
 \caption{As above but for $M_J, (J-K)_0$ colours. Here the effects
   of high accretion can shift the stars into higher or lower mass bins.
   \label{high_acc_ir}}
\end{figure*}

As can be seen by comparing Figure \ref{spread_ir_JJK} and Figure
\ref{high_acc_ir}, accreting BD systems without discs are shifted
brighter and bluer resulting in a large error in isochronal mass
derivation. The simulated photometry for BDD systems however, shows
large magnitudes of scatter redward and towards slightly brighter
magnitudes. Given that the reddening vector is approximately
horizontal in a \textit{J, J-K} CMD many of these objects will appear
as heavily obscured objects of higher mass. The \textit{top left
  panel} shows that the scatter does increase for higher accretion
rates, however, even at the lower of the accretion rates (-8
log$(\frac{\dot{M}}{M_{\odot}}yr^{-1})$), significant scatter is
evident. The \textit{top right panel} shows that the scatter is
roughly the same for both areal coverages. The \textit{bottom left
  panel} shows, again, no correlation between rotation rate and
scatter within the CMD. This is unsurprising as at these elvated
accretion rates the increase in flux will lead to significant levels
of dust sublimation of the inner edge, as discussed in more detail in
Section \ref{disc_struct}. Finally, the \textit{bottom right panel}
shows systems with a larger inclination having, generally lower
magnitudes, and therefore lower masses, as expected.

Therefore, again as for the optical case, given our assumptions and
parameter space, BDD systems accreting at higher rates would not be
identified with the correct mass in most cases using IR CMDs.

\subsubsection{Disc fractions}
\label{derive_frac}

To distinguish between the proposed formation mechanisms for BDs, and
to derive disc lifetimes, accurate derivations of disc fractions are
required (as well as stellar masses).

\textbf{Derivation of disc fractions.}

>>>>

Data for rotation rates, from periodic variability surveys, are widely
available for a range of different age clusters of TTS. However, fewer
studies exist on the rotation rates of BD mass objects. To select a
feasible range of periods for our model grid we have collated the
results from some recent studies of BD variability. Rotation period
data for $\sigma$ Ori, at an age of $\approx$ 3 Myrs
\citep{2008MNRAS.386..261M}, is studied in \cite{2004A&A...419..249S},
where periods are found over the range 5.78$-$74.4 hours ($\approx$
0.24$-$3.1 days) for BD mass objects. In some cases the periodic
variability is irregular and assumed to come from active accretion hot
spots on the BD surface \citep[see][for discussion of variability
causes]{1995A&A...299...89B,2007prpl.conf..297H}. Infrared excesses
and evidence of accretion are also found to correlate with slower
rotating TTS, which the authors construe as evidence for a disc
locking scenario. \cite{2004A&A...419..249S} also find a $mass\propto
period$ relationship extending into the BD mass regime, this faster
rotation in younger mass objects is taken as evidence for a decreasing
effectiveness of the angular momentum removal of the disc locking
mechanism, i.e. imperfect disc locking. In addition,
\cite{2004A&A...419..249S} argue that when compared with data from
younger clusters disc lifetimes can be seen to decrease with central
object mass.

\cite{2005A&A...429.1007S} study rotation period data for stars in the
vicinity of $\epsilon$ Ori, with an assumed age of $\approx$3 Myrs
\citep{2002A&A...384..937Z} and the ONC at an age of $\approx$2 Myrs
\citep{2008MNRAS.386..261M}. Rotation periods, from photometric
variability, in the range 4.7$-$87.6 hours ($\approx$ 0.2$-$3.65 days)
for BD mass stars are found. Again, \cite{2005A&A...429.1007S} find
some BD objects with irregular variability indicative of accretion hot
spots on the stellar surface. \cite{2005A&A...429.1007S} then analyse
the mass-period relationship of these data in comparison with data
from $\sigma$ Ori \citep[3 Myr,][]{2004A&A...419..249S}, NGC2264
\citep[2 Myr,][]{2004A&A...417..557L} and then ONC \citep[1
Myr][]{2002A&A...396..513H}. They conclude that either the $\epsilon$
Ori sample has an age of around 2$-$3 Myrs or there is still
significant disc locking (concluded using the slope of the mass-period
relation). More specifically for the BD mass regime, they find
\citep[in agreement with][]{2004A&A...419..249S} an age independent BD
lower period limit of 2$-$4 hours, therefore at any age a fraction of
BD stars rotate at short and constant period. Given that these objects
should spin up as they hydrostatically contract they invoke rotational
braking in the BD mass regime as a mechanism of removing significant
angular momentum.

\cite{2003ApJ...594..971J} study the rotational periods of BD (and
very low mass stars) in the Chameleon I region. This region is $\le$ 1
Myrs old, and the authors find rotation periods of 2.19, 3.376 and
3.21 days for their BD mass counterparts. Using supplementary data
from the literature \cite{2003ApJ...594..971J} propose that a shorter
lifetime of $\approx$5 Myrs for accretions discs around BDs is more
probable and is inferred from the derivation of a shorter locking
timescale. However, as discussed in the previous studies, an imperfect
disc locking mechanisms is also hypothesised as responsible for the
less significant loss in angular momentum out to ages of 10 Myrs. The
data mentioned above and other datasets are summarised and discussed
in \cite{2007prpl.conf..297H}, where a strong argument for the
mass-period relationship can be found, shown in their Fig 6, alongside
further support for either a shorter disc lifetime or imperfect disc
locking mechanism for BD stars.

Therefore, in summary, it is currently unclear if any disc locking
mechanisms efficiency or disc lifetimes are a function of mass, with
an associated phase change over the TTS to BD mass limit. Rotation
rate and disc presence data are being used to explore issue.
Therefore, to create a set of useful models to help contextualise the
observational constraints we must adopt a realistic, and bounding,
range of rotation rates. For our model grid, and associated age range
($<$10 Myrs), we have selected 0.5 and 5 days.  The lower limit is
placed at around the approximate median for the faster rotating
samples, providing a useful benchmark for exploring possible detection
limits of faster rotators with discs. The longer period is set at the
approximate edge of the slow rotating tail, providing a limiting case
for the slower rotating, disc-locked, candidates.
